Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}

### 2021-07-07 Latex/sections/realData.tex
LPC results added, rolling sphere added.

We also tested our algorithm on a real-world dataset that adheres to different motion profiles and laser scanners type with different scanning patterns than the simulated data.

The dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
On this floating desk, the sphere can rotate freely about all axis while being fixed in position. 
Hence, in this experiment the optimization space can be reduced to a rotation.
Further, without the motion, the 2D laser cannot obtain a 3D model from the environment, thus requiring registration. 
The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the orientation, a battery pack and a raspberry pie for processing.
This IMU provides a precision of \SI{76.3}{\micro g} in linear acceleration, \SI{0.02}{\degree\per\second} in angular velocity about the x and y axis, and \SI{0.013}{\degree\per\second} in angular velocity about the z axis \cite{PhidgetSpatial}.

Figure~\ref{fig:float-sat-sphere} shows the experimental setup. The left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud. 
We see that the point cloud is rather noisy, and in particular, the walls are sensed multiple times, hence appearing very blurry.

@@

>>>>

We also tested our algorithm on two real-world datasets that adhere to different motion profiles and laser scanner types with different scanning patterns.

The first dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
On this floating desk, the sphere can rotate freely about all axis while being fixed in position. 
Hence, in this experiment the optimization space can be reduced to a rotation.
Further, without the motion, the 2D laser cannot obtain a 3D model from the environment, thus requiring registration. 
The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the orientation, a battery pack and a raspberry pie for processing.
This IMU provides a precision of \SI{76.3}{\micro g} in linear acceleration, \SI{0.02}{\degree\per\second} in angular velocity about the x and y axis, and \SI{0.013}{\degree\per\second} in angular velocity about the z axis \cite{PhidgetSpatial}.  
Figure~\ref{fig:float-sat-sphere} shows the experimental setup. The left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud. 
We see that the point cloud is rather noisy, and in particular, the walls are sensed multiple times, hence appearing very blurry.

@@

The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor that the simulation uses.
We put the laserscanner into an acrylic glass spherical shell together with three IMUs and roll it by hand with a slow, constant motion in a home indoor environment.
Thus, this time almost all six degrees of freedom are used for optimization.
We exlude a translation in the global up- or downwards direction, though, since we also locked that dimension when collecting the data.
Figure~\ref{fig:prototype} shows the prototpye.
The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the pure transformation found by the recorded pose.
Although the robot has been moved carefully slow, you see that pose error accumulates in such a way that the same wall is visible multiple times.

\begin{figure}
  \includegraphics[width=0.49\linewidth]{./images/photoReal1.jpeg}
  \includegraphics[width=0.50\linewidth]{./images/photoReal3.jpeg}
  \caption{Prototype used for data acquisition with the rolling sphere. The main payload is the Livox Mid-100 laser-scanner. For pose-estimation, three IMUs of the manufacturer Phidget are placed inside and a Raspberry Pi 4 for the calculations. On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner. }
  \label{fig:prototype}
\end{figure}

### 2021-07-07 Latex/sections/registration.tex
LPC results added, rolling sphere added.

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as the square of the L2-norm of the distance:
\begin{align}
	E = \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_k} \sum_{\vec{p}_i\in \mathcal{P}_k}2D(\vec{\Pi}, \vec{p}_i)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{align}
    &\begin{alignedat}{3}
        \nabla E_\varphi = &n_{\mathcal{P}_k}^y&&(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] \\&&&- z_i C_\varphi C_\vartheta)\\
        &+ n_{\mathcal{P}_k}^z&&(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] \\&&&- z_i S_\varphi C_\vartheta)
    \end{alignedat} \\
    &\begin{alignedat}{3}
        \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
        &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi S_\vartheta)
    \end{alignedat}\\
    &\begin{alignedat}{3}
       \nabla E_\psi =& n_{\mathcal{P}_k}^x&&\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
       &+ n_{\mathcal{P}_k}^y&&(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] \\
       &&&+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
       &+ n_{\mathcal{P}_k}^z&&(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
       &&&+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
    \end{alignedat}
\end{align}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

>>>>

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as the square of the L2-norm of the distance:
\begin{align}
	E = \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_k} \sum_{\vec{p}_i\in \mathcal{P}_k}2D(\vec{\Pi}, \vec{p}_i)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{equation}
\begin{split}
\nabla E_\varphi = n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]+y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] - z_i C_\varphi C_\vartheta)\\
        + n_{\mathcal{P}_k}^z(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] +y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
\end{split} \;\; ,
\end{equation}
\begin{equation}
\begin{split}
 \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
        &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi 
\end{split}
\;\; ,
\end{equation} 
\begin{equation}
\begin{split}
	\nabla E_\psi = &n_{\mathcal{P}_k}^x\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
       &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] 
       + y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
       &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] 
       + y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
\end{split}
\end{equation}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

### 2021-07-07 Latex/sections/results.tex
LPC results added, rolling sphere added.

Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
The pre-registration is obtained by determining the orientation of the sphere via Madgwick filtered~\cite{madgwick2010efficient} IMU measurements.
To increase the number of possible point-to-plane correspondences we combine ten temporally successive scans into one meta-scan, which is then globally registered. 
For this we must assume constant pose over this time duration. 
However, given the high measurement frequency of the Lidar this introduces only a small error. 
This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
Transformations like these happen in particular for a small collection of points as outliers have more influence.

\begin{figure*}
	\centering
	\includegraphics[width=0.495\textwidth]{./images/cylon_uncorr_corner}\hfill
	\includegraphics[width=0.495\textwidth]{./images/cylon_corr_corner}\\
	\includegraphics[width=0.495\textwidth]{./images/cylon_uncorr_top}\hfill
	\includegraphics[width=0.495\textwidth]{./images/cylon_corr_top}
	\caption{The point cloud acquired by the floating sphere before (left) and after (right) applying the plane based registration. View from the interior that shows the used planes (top) and a birds-eye view (bottom). An animation of the registration process is given at \url{https://youtu.be/8XdIUN_9VpY} .}
	\label{fig:cylon-corrected}
\end{figure*}
After the registration, the walls of the room are significantly more prominent in the point cloud. 
In particular, the noise in the top left corner of the top view (cf. figure~\ref{fig:cylon-corrected}) has been visibly reduced.
Further, the deviation of points around the walls is notably smaller as the points are moved on the plane.

However, some imprecisions remain. 
In particular, at the initially very noisy top left corner of the top view, a few points cannot be matched to the plane. 
The scans that cause these imprecisions are problematic as they measure a number of plane-like structures in the interior of the room. 
When relabeling those scans, the interior points are associated with a plane, and hence the corresponding distance is minimized.

>>>>

Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
The pre-registration is obtained by determining the orientation of the sphere via Madgwick filtered~\cite{madgwick2010efficient} IMU measurements.
To increase the number of possible point-to-plane correspondences we combine twenty temporally successive scans into one meta-scan, which is then globally registered. 
We always use the middlest line scan as a reference coordinate system.
This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
Transformations like these happen in particular for a small collection of points as outliers have more influence.

\begin{figure*}
	\centering
	\includegraphics[width=0.495\textwidth]{./images/cylon_uncorr_corner}\hfill
	\includegraphics[width=0.495\textwidth]{./images/cylon_corr_corner}\\
	\includegraphics[width=0.495\textwidth]{./images/cylon_uncorr_top}\hfill
	\includegraphics[width=0.495\textwidth]{./images/cylon_corr_top}
	\caption{The point cloud acquired by the floating sphere before (left) and after (right) applying the plane based registration. View from the interior (top) and a birds-eye view (bottom). Parameters used for optimization: $\epsilon_H = 50$, $\epsilon_P = 200$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 20$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 200$, $\alpha_0 = [0.01, 0.01, 0.01, 0, 0, 0]^\tau$, $i = 8$, $k = 100$. An animation of the registration process is given at \url{https://youtu.be/Qa2Qi8z0KEk} .}
	\label{fig:cylon-corrected}
\end{figure*}
After the registration, the walls of the room are significantly more prominent in the point cloud. 
In particular, the noise in the top left corner of the top view (cf. figure~\ref{fig:cylon-corrected}) has been visibly reduced.
Further, the deviation of points around the walls is notably smaller as the points are moved on the plane.

\subsection{Rolling Sphere Results}

Figure~\ref{fig:jasperhome} shows the results obtained before and after employing the presented algorithm on the dataset, acquired by the prototype in figure~\ref{fig:prototype}.
We again combine twenty temporally successive scans into one meta-scan, using the middlest scan as a reference coordinate system, which is then globally registered.

\begin{figure*}
        \centering
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_corner}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_corner}\\
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_top}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_top}
        \caption{The point cloud acquired by the rolling sphere before (left) and after (right) applying the plane based registration. View from the interior (top) and a birds-eye view (bottom). Parameters used for optimization: $\epsilon_H = 200$, $\epsilon_P = 300$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 200$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 800$, $\alpha_0 = [0.001, 0.001, 0.001, 1, 0, 1]^\tau$, $i = 1500$, $k = 1$. An animation of the registration process is given at \url{https://youtu.be/hYR3xLhIdsE}.}
        \label{fig:jasperhome}
\end{figure*}
After the registration, the walls of the environment are no longer ambiguous, since all the sub-sequent scans got matched with the initial plane representation.
However, the improved path looks distorted.
This is because for some scans, especially empty ones where the sensor points to the ground, no planar features got matched with the global plane model, thus their pose is not optimized.
Improving the path quality is an objective for future works.
